The contact value approximation for the pair distribution function of the inhomogeneous hard-sphere fluid

### 2014-06-17 papers/pair-correlation/paper.tex
help Dan get his slides sconsing Tested on bennet. Took 1:00:04.
edit introduction to refer to g(n) vs g(V) Tested on bennet. Took 0:21.51.

At its core, thermodynamic perturbation theory---sometimes referred to
as the cluster, or high-temperature expansion---is an expansion of the
free energy in powers of a small parameter, which is typically a
pairwise attractive interaction:
\begin{align}
  F &= F_0 + F_1 + \beta F_2 + \mathcal{O}(\beta^2)
\end{align}
where the terms $F_n$ are corrections to the free energy of order $n$
in the small interaction.  The first and largest term in this
expansion is
\begin{align}
  F_1 &= \tfrac12 \iint
  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
  d\rr_1d\rr_2
  \label{eq:mean-field}
\end{align}
where $g^{(2)}_{HS}(\rr_1,\rr_2)$ is the pair distribution function of
the hard-sphere reference fluid, and $\Phi(r)$ is the pair potential.
In this paper, we introduce a contact value approximation (CVA) for
the hard-sphere pair distribution function which is suitable for use
in the creation of classical density functionals based on
thermodynamic perturbation theory. We then demonstrate a more
computationally efficient separable version of the contact
value approximation (CVA-S), which is based on a fit to the radial
distribution function that is separable in a way that will enable
efficient evaluation of the integral in Eq.~\ref{eq:mean-field}.

\fixme{Somewhere above introduce the question of whether g2 is
  expressed in terms of V or n}\davidsays{I'll take a stab at this}

\section{Previous theoretical approaches $g^{(2)}(\rr_1,\rr_2)[V(\rr)]$}

@@

\section{Previous approximations to the pair distribution function
  $g^{(2)}(\rr_1,\rr_2)[n(\rr)]$}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
  \caption{Plot of the hard-sphere radial distribution function, with
    our separable fit.  The solid lines show the true radial
    distribution function $g(r)$, as found from Monte Carlo
    simulation, and the dashed lines show our separable fit to the
    radial distribution functions
    $g_S(r)$.}\label{fig:radial-distribution}
\end{figure}

>>>>

At its core, thermodynamic perturbation theory---sometimes referred to
as the cluster, or high-temperature expansion---is an expansion of the
free energy in powers of a small parameter, which is typically a
pairwise attractive interaction:
\begin{align}
  F &= F_0 + F_1 + \beta F_2 + \mathcal{O}(\beta^2)
\end{align}
where the terms $F_n$ are corrections to the free energy of order $n$
in the small interaction.  The first and largest term in this
expansion is
\begin{align}
  F_1[n(\rr)] &= \tfrac12 \iint \!\!
  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
  d\rr_1d\rr_2
  \label{eq:mean-field}
\end{align}
where $g^{(2)}_{HS}(\rr_1,\rr_2)$ is the pair distribution function of
the hard-sphere reference fluid, and $\Phi(r)$ is the pair potential.
Formally, this requires the pair distribution function as a functional
of the density $n(\rr)$.  In Section~\ref{sec:gV}, we will introduce
existing theoretical approaches for computing
$g^{(2)}_{HS}(\rr_1,\rr_2)$ given the external potential felt by the
hard spheres.  In Section~\ref{sec:gn}, we will introduce existing
approximations for the hard-sphere pair distribution that are
expressed as a functional of the density distribution $n(\rr)$, which
is a form that is more directly useful in the construction of
classical density functionals---which are themselves expressed as a
functional of the density.

In this paper, we introduce a new contact value approximation (CVA) for
the hard-sphere pair distribution function which is suitable for use
in the creation of classical density functionals based on
thermodynamic perturbation theory. We then demonstrate a more
computationally efficient separable version of the contact
value approximation (CVA-S), which is based on a fit to the radial
distribution function that is separable in a way that will enable
efficient evaluation of the integral in Eq.~\ref{eq:mean-field}.

\section{Previous theoretical approaches $g^{(2)}(\rr_1,\rr_2)[V(\rr)]$}\label{sec:gV}

@@

\section{Previous approximations to the pair distribution function
  $g^{(2)}(\rr_1,\rr_2)[n(\rr)]$}\label{sec:gn}

@@

\begin{figure}
  \centering
  \includegraphics[width=\columnwidth]{figs/ghs-g2}
  \caption{Plot of the hard-sphere radial distribution function, with
    our separable fit.  The solid lines show the true radial
    distribution function $g(r)$, as found from Monte Carlo
    simulation, and the dashed lines show our separable fit to the
    radial distribution functions
    $g_S(r)$.}\label{fig:radial-distribution}
\end{figure}
